---
title: "Data process"
output: 
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load the dataset we want.
```{r}
library(tidyverse)
library(janitor)
library(ggridges)
library(ggthemes)


library(stringr)
library(dplyr)
library(forcats)

#embedding plots in rmarkdown
knitr::opts_chunk$set(fig.width=12, fig.height=8, out.width = "80%")
theme_set(theme_bw())
```


First we import the data and clean it.
```{r}
health = 
  readxl::read_xls('../food_enviroment_atlas.xls', sheet = 'HEALTH') %>%
  clean_names() %>%
  select(1:7)

socioeconomic = 
  readxl::read_xls('../food_enviroment_atlas.xls', sheet = 'SOCIOECONOMIC') %>%
  clean_names() %>%
  select(1:3, 10:18)

assistance = 
  readxl::read_xls('../food_enviroment_atlas.xls', sheet = 'ASSISTANCE') %>%
  clean_names() %>%
  select(1:3, 23:29)

restaurant = 
  readxl::read_xls('../food_enviroment_atlas.xls', sheet = 'RESTAURANTS') %>%
  clean_names() %>%
  select(1:9, 16:17)

county =
  readxl::read_xls('../food_enviroment_atlas.xls', sheet = 'Supplemental Data - County') %>%
  clean_names()

state =
  readxl::read_xls('../food_enviroment_atlas.xls', sheet = 'Supplemental Data - State') %>%
  clean_names() %>%
  select(1:2, 9:20, 27:40)

store = 
  readxl::read_xls('../food_enviroment_atlas.xls', sheet = 'STORES') %>%
  clean_names() %>%
  select(1:27)

# median household income
medhincome_08 = readxl::read_xls("../mhi_08.xls", range = "A13:B63",col_names = c("state","medhhinc08")) %>%
  clean_names() %>%
  mutate(state = state.abb[match(state,state.name)])
medhincome_08$state[17] = "DC"
medhincome_13 = readxl::read_xls("../mhi_13.xls", range = "A11:B61",col_names = c("state","medhhinc13")) %>%
  clean_names() %>%
  mutate(state = state.abb[match(state,state.name)])
medhincome_13$state[8] = "DC"
medhincome_13$`income level` = 
  ifelse(medhincome_13$medhhinc13 < 47242.68,"below average",
         ifelse(medhincome_13$medhhinc13 < 58380.28, "middle income", "above average"))
medhincome = merge(medhincome_08, medhincome_13,by=c("state"))
```

### Distribution of obeses and diabetes
We first take a look at the distribution of diabetes and obesity in USA in 2013. Deeper color indicates higher rates of diabety or obetes. Southeastern part has higher diabetes and obesity rates. Especially, Alamba has the highest diabetes and obesity rate. 
```{r}
library(plotly)
health_map = 
  health %>%
  group_by(state) %>%
  summarise(diabetes = mean(pct_diabetes_adults13),
            obesity = mean(pct_obese_adults13)) %>%
  ungroup() %>%
  na.omit() %>%
  mutate(full_state = state.name[match(state, state.abb)],
         hover = with(., paste(full_state, '<br>')))

geo_info <- list(
  scope = 'usa',
  projection = list(type = 'albers usa'),
  showlakes = TRUE,
  lakecolor = toRGB('white')
)

plot_geo(health_map, locationmode = 'USA-states') %>%
add_trace(
  z = ~diabetes, text = ~hover, locations = ~state,
  color = ~diabetes, colors = 'OrRd'
) %>%
colorbar(title = "Diabetes Rate (%)") %>%
layout(
  title = 'Adult Diabetes Rate, 2013',
  geo = geo_info
  )


plot_geo(health_map, locationmode = 'USA-states') %>%
add_trace(
  z = ~obesity, text = ~hover, locations = ~state,
  color = ~obesity, colors = 'YlGnBu'
) %>%
colorbar(title = "Obesity Rate (%)") %>%
layout(
  title = 'Adult Obesity Rate, 2013',
  geo = geo_info
  )

```

We then take a look at the distribution of diabetes and obesity nationally.  
Paired t-test results indicates that both diabetes and obesity rate significantlly increase from 2008 to 2013. This conclusion is consist with our literature research. Diabtes and obesity rates keep rising in the past decades in USA.(Mokdad A H, et al., 2001)
```{r}
require(gridExtra)

t_test_diabetes = 
  health %>%
  ggplot() +
  geom_density(aes(x = pct_obese_adults08), color = "#56B4E9", fill ="#56B4E9", alpha = .4) +
  geom_density(aes(x = pct_obese_adults13), color = "#D55E00",fill ="#D55E00", alpha = .4) +
  labs(title = 'Comparisons between the rate of obesity in 2008 and 2013',
       x = 'rate of obesity')

t.test(health$pct_obese_adults08, health$pct_obese_adults13, paired=TRUE)

t_test_obesity = 
  health %>%
  ggplot() +
  geom_density(aes(x = pct_diabetes_adults08), color = "#56B4E9", fill ="#56B4E9", alpha = .4) +
  geom_density(aes(x = pct_diabetes_adults13), color = "#D55E00",fill ="#D55E00", alpha = .4) +
  labs(title = 'Comparisons between the rate of obesity in 2008 and 2013',
       x = 'rate of diabetes')

t.test(health$pct_obese_adults08, health$pct_obese_adults13, paired = TRUE)
t.test(health$pct_diabetes_adults08, health$pct_diabetes_adults13, paired = TRUE)

grid.arrange(t_test_diabetes, t_test_obesity, ncol=2)

```

Finding the similar distribution of diabetes and obesity, we continue to test the relationship between diabetes and obesity in 2008 and 2013 respectively. After the linear regression, we find that there are significant linear relationship between diabetes and obesity both in 2008 and 2013. Around 50% variation of diabetes could be explained by obesity. This conclusion is consistent with other national researches.(Nguyen, et al., 2011)
```{r}
require(gridExtra)

relation_2008 = 
  health %>%
  ggplot() +
  geom_point(aes(x = pct_obese_adults08, y = pct_diabetes_adults08), color="#56B4E9")+
  geom_smooth(aes(x = pct_obese_adults08, y = pct_diabetes_adults08), color="#56B4E9", method = 'lm') +
  labs(title = 'Relationship between the obesity and diabetes rate in 2008',
       x = 'obesity percentage',
       y = 'diabetes percentage')

relation_2013 = 
  health %>%
  ggplot() +
  geom_point(aes(x = pct_obese_adults13, y = pct_diabetes_adults13), color="#D55E00")+
  geom_smooth(aes(x = pct_obese_adults13, y = pct_diabetes_adults13), color="#D55E00", method = 'lm') +
  labs(title = 'Relationship between the obesity and diabetes rate in 2013',
       x = 'obesity percentage',
       y = 'diabetes percentage')
grid.arrange(relation_2008, relation_2013, ncol=2)

summary(lm(pct_diabetes_adults08~pct_obese_adults08, data = health))
summary(lm(pct_diabetes_adults13~pct_obese_adults13, data = health))
```

### Socioeconomic VS diabetes and obesity
```{r}
# extract relavent data
diabetes = health %>%
  group_by(state) %>%
  summarise(pct_08 = mean(pct_diabetes_adults08),
            pct_13 = mean(pct_diabetes_adults13)) %>%
  gather(key = year, value = pct_diabetes_adults, pct_08:pct_13)%>%
  separate(year, into = c("remove", "year"), sep = "_") %>%
  select(-remove)

obese = health %>%
  group_by(state) %>%
  summarise(pct_08 = mean(pct_obese_adults08),
            pct_13 = mean(pct_obese_adults13)) %>%
  gather(key = year, value = pct_obese_adults, pct_08:pct_13)%>%
  separate(year, into = c("remove", "year"), sep = "_") %>%
  select(-remove)

health_status = merge(diabetes, obese, by=c("state", "year"))

income = medhincome %>%
  gather(key = year, value = medhhinc, medhhinc08:medhhinc13)%>%
  separate(year, into = c("remove", "year"), sep = -3) %>%
  select(-remove)

eco_health = merge(health_status, income, by=c("state", "year"))
```


```{r}
# plots
# income VS obesity
eco_health %>%
  group_by(year) %>%
  ggplot(aes(x = medhhinc, y = pct_obese_adults)) +
  geom_point(aes(color = year), size = 3, alpha = .8) +
  geom_smooth(se = FALSE) + 
  facet_grid(. ~ year) +
  labs(
    x = "Median household income",
    y = "Percentage of adult obesity")  +  
  theme(text = element_text(size = 16), 
        axis.text.x = element_text(size = 14), 
        axis.text.y = element_text(size = 14))

# income VS diabetes
eco_health %>%
  group_by(year) %>%
  ggplot(aes(x = medhhinc, y = pct_diabetes_adults)) +
  geom_point(aes(color = year), size = 3, alpha = .8) +
  geom_smooth(se = FALSE) + 
  facet_grid(. ~ year) +
  labs(
    x = "Median household income",
    y = "Percentage of adult obesity")  +  
  theme(text = element_text(size = 16), 
        axis.text.x = element_text(size = 14), 
        axis.text.y = element_text(size = 14))

```


### Health related figure
* rate of diabetes increases
```{r}
health %>%
  ggplot() +
  geom_density(aes(x = pct_diabetes_adults08), fill="#56B4E9") +
  geom_density(aes(x = pct_diabetes_adults13), fill="#D55E00") +
  labs(title = 'Comparisons between the rate of diabetes in 2008 and 2013',
       x = 'rate of dabetes')

t.test(health$pct_diabetes_adults08, health$pct_diabetes_adults13)
```

* rate of obesity increases
```{r}
health %>%
  ggplot() +
  geom_density(aes(x = pct_obese_adults08), fill="#56B4E9") +
  geom_density(aes(x = pct_obese_adults13), fill="#D55E00") +
  labs(title = 'Comparisons between the rate of obesity in 2008 and 2013',
       x = 'rate of obesity')

t.test(health$pct_obese_adults08, health$pct_obese_adults13)
```

* Linear relationship between rate of diabety and obesity
```{r}
require(gridExtra)
relation_2008 = 
  health %>%
  ggplot() +
  geom_point(aes(x = pct_diabetes_adults08, y = pct_obese_adults08), color="#56B4E9")+
  geom_smooth(aes(x = pct_diabetes_adults08, y = pct_obese_adults08), color="#56B4E9", method = 'lm') +
  labs(title = 'Relationship between the diabetes rate and obesity rate in 2008',
       x = 'diabetes percentage',
       y = 'obesity percentage')
relation_2013 = 
  health %>%
  ggplot() +
  geom_point(aes(x = pct_diabetes_adults13, y = pct_obese_adults13), color="#D55E00")+
  geom_smooth(aes(x = pct_diabetes_adults13, y = pct_obese_adults13), color="#D55E00", method = 'lm') +
  labs(title = 'Relationship between the diabetes rate and obesity rate in 2013',
       x = 'diabetes percentage',
       y = 'obesity percentage')
grid.arrange(relation_2008, relation_2013, ncol=2)
# summary(lm(pct_diabetes_adults08~pct_obese_adults08, data = health))
```
Here we find that there are linear relationship between diabetes and obesity.

* Compared poverty rate in 2015 with the obesity rate in 2013
```{r}
poverty_health = 
  left_join(health, socioeconomic, by = "fips")

poverty_health %>%
  ggplot(aes(x = povrate15, y = pct_obese_adults13)) +
  geom_point(color = "#56B4E9") +
  labs(title = 'Compared poverty rate in 2015 with the obesity rate in 2013',
       x = 'poverty in 2015',
       y = 'obesity rate in 2013')
```
It seems that data are clustered.  

### Lunch Program in each state
```{r}
state_population = state[-52,] %>% 
  gather(key = "year1", value = "population",
         state_population_2009:state_population_2016) %>% 
  mutate(year1 = str_replace(year1, "state_population_","")) %>% 
  select(state, year1, population) %>% 
  rename(state1 = state) %>% 
  filter(!(year1 == "2010" | year1 == "2016"))

#prepare dataset for lunch program
lunch = state[-52,] %>% 
  gather(key = "year", value = "lunch_participants", national_school_lunch_program_participants_fy_2009:national_school_lunch_program_participants_fy_2015) %>% 
  mutate(year = str_replace(year, "national_school_lunch_program_participants_fy_", "")) %>% 
  select(statefips, state, year, lunch_participants) %>% 
  left_join(state_population, by = c("year" = "year1", "state" = "state1")) %>% 
  mutate(lunch_proportion = lunch_participants/population)

breakfast = state[-52,] %>% 
  gather(key = "year", value = "breakfast_participants",
         school_breakfast_program_participants_fy_2009:
         school_breakfast_program_participants_fy_2015) %>% 
  mutate(year = str_replace(year, "school_breakfast_program_participants_fy_",""),
         breakfast_participants = as.numeric(breakfast_participants)) %>% 
  select(statefips, state, year, breakfast_participants) %>% 
  left_join(state_population, by = c("year" = "year1", "state" = "state1")) %>% 
  mutate(breakfast_proportion = breakfast_participants/population)

summer = state[-52,] %>% 
  gather(key = "year", value = "summer_participants",
         summer_food_particpants_fy_2009:
         summer_food_participants_fy_2015) %>% 
  mutate(year = str_replace(year, "summer_food_particpants_fy_","")) %>% 
  mutate(year = str_replace(year, "summer_food_participants_fy_","")) %>% 
  select(statefips, state, year, summer_participants) %>% 
  left_join(state_population, by = c("year" = "year1", "state" = "state1")) %>% 
  mutate(summer_proportion = summer_participants/population)
```

### The relationship between lunch program and diabetes
```{r}
lunch_spread = lunch %>% 
  select(-lunch_participants,-population,-state) %>% 
  rename(year1 = year, statefips1 = statefips) %>% 
  spread(key = "year1", value = "lunch_proportion")

breakfast_spread = breakfast %>% 
  select(-breakfast_participants,-population,-state) %>% 
  rename(year1 = year, statefips1 = statefips) %>% 
  spread(key = "year1", value = "breakfast_proportion")

summer_spread = summer %>% 
  select(-summer_participants,-population,-state) %>% 
  rename(year1 = year, statefips1 = statefips) %>% 
  spread(key = "year1", value = "summer_proportion")

health_13 = health %>% 
  group_by(state) %>% 
  mutate(diabetes_13 = mean(pct_diabetes_adults13, na.rm = T),
         obesites_13 = mean(pct_obese_adults13, na.rm = T)) %>% 
  mutate(statefips = str_sub(fips, 1, 2))%>% 
  filter(!duplicated(state)) %>% 
  ungroup(state) %>%
  select(statefips, state, diabetes_13, obesites_13) %>% 
  left_join(lunch_spread[,c(1,5)], by = c("statefips" = "statefips1")) %>% 
  rename("lunch" = `2013`) %>% 
  left_join(breakfast_spread[,c(1,5)], by = c("statefips" = "statefips1")) %>% 
  rename("breakfast" = `2013`) %>% 
  left_join(summer_spread[,c(1,5)], by = c("statefips" = "statefips1")) %>% 
  rename("summer" = `2013`) %>% 
  left_join(medhincome_13, by = "state") %>% 
  gather(key = "program", value = "participants proportion", lunch:summer)

#linear regression: diabetes_13 v.s. lunch participants (2013)
#diabetes_13 = 22.991 + 5.235*ln(lunch_participants_proportion)
#diabetes_13 = e^(22.991)*(e(5.235))^(lunch_participants_proportion )
ggplot(health_13,aes(x = log(`participants proportion`), y = diabetes_13)) +
  geom_point(aes(color = `income level`)) +
  geom_smooth(method = "lm",formula = y~x) +
  coord_cartesian(xlim = c(-6, -2)) +
  labs(title = "Diabetes Rate ~ ln(partic-2pation rate)",
    x = "",
    y = "Diabetes Rate") +
  theme(text = element_text(size = 15),
        legend.position = "bottom") +
  facet_grid(. ~program)
```
