---
title: "Diabetes Rate and Food Environment"
author: "Shengnan Sun(ss5428), Shumin Rui(sr3461), Jian Zou(jz2924), Hsiang-Lun Kao(hk2968)"
output: 
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

library(tidyverse)
library(janitor)
library(ggridges)
library(ggthemes)
library(stringr)
library(dplyr)
library(forcats)
library(plotly)

theme_set(theme_bw())
theme_update(legend.position = "bottom")
```

## Introduction
Diabetes is a chronic disease whose prevalence has increased rapidly in the past few decades. According to the report from WHO, in 2014, the prevalence of diabetes among adult population had risen from 4.7% to 8.5%  comparing to that in 1980 (WHO, 2016). It is projected as the seventh leading cause of mortality in 2030 (Mathers & Loncar, 2006).  

In the U.S., the data of 2015 shows that 7.2% of the population are diagnosed with Diabetes(CDC, 2017). In 2013, it ranked as the seventh leading cause of death in the nation (CDC). Previous studies identified obesity as the main risk factor of type 2 diabetes (which account for 95% cases of diabetes in the U.S.)(Smyth & Heron, 2006). Meanwhile,  consumption of artificial sweetened beverage(Koning et al, 2011) and  prenatal exposure to famine under certain circumstances(Lumey, Khalangot & Vaiserman, 2015) are both related to the increased prevalence of type 2 diabetes. The report from CDC also demonstrates that there are certain association between socioeconomic status and the prevalence(CDC, 2017). 

Inspiring by these findings which are all somehow related to food accessibility, we set the goal to explore the association between food related factors, including median household income, participation proportion of National School Lunch Program (NSLP)/School Breakfast Program (SBP), Fast-food restaurants availability, etc. and the prevalence of diabetes in the US. Over the course of the research, we further want to build a model to explain the relationship between food related factors and the prevalence of Diabetes.

## Data and Method
The dataset we would like to use is <B>Food Environment Atlas</B> which can be found on [data.gov](https://catalog.data.gov/dataset/food-environment-atlas-f4a22).The detailed description of the dataset can be found on the [Github page](https://github.com/jianzou95/DS-Final-Project/blob/master/archived_documentation_August2015.pdf) of the project. The detailed description of the dataset can be found on the Github page of the project, named archived_documentation_August2015.pdf. The dataset contains several spreadsheets which provide information about population characteristics and food factors of each state at the county level.  Some main variables that we are interested in includ:

* Adult diabetes and obesity rate in every countys in 2008 and 2013. They have been summarized into states by evening across counties.  
* Participation percentage of National School Lunch Program and School Breakfast Program in 2009 and 2014.
* The number of Fast-food restaurants in 2014, the number of Fast-food restaurants per 1,000 population in  2014 and the Expenditures per capita of fast food in 2012. 


In addition to this main dataset, we also use another two dataset, <B>Median Household Income (In 2013 Inflation-adjusted Dollars)</B> by [State Ranked](https://www.census.gov/data/tables/2014/demo/income-poverty/p60-249.html) from Highest to Lowest Using 3-Year Average (2011-2013), and  <B>Median Household Income (In 2008 Inflation-adjusted Dollars)</B> by [State Ranked](https://www.census.gov/data/tables/2009/demo/income-poverty/p60-236.html.) from Highest to Lowest Using 3-Year Average: 2006-2008 from census.gov. 

The two median household income datasets contain the information of median household income of each state (50 states and DC) in 2008 and 2013 collected by U.S. Census Bureau. The income values are evened in a three-year-period (ie. 2006-2008 for 2008, and 2011-2013 for 2013) and adjusted for inflation. 

* These two datasets have been merged into one dataframe based on states, and the names of states are converted to abbreviation in order to be compared with the rest dataframes.

All datasets are downloaded directly from the websites as Excel format. 

```{r data_import}
health = 
  readxl::read_xls('../food_enviroment_atlas.xls', sheet = 'HEALTH') %>%
  clean_names() %>%
  select(1:7)

assistance = 
  readxl::read_xls('../food_enviroment_atlas.xls', sheet = 'ASSISTANCE') %>%
  clean_names() %>%
  select(1:3, 23:24, 30:31)

restaurant = 
  readxl::read_xls('../food_enviroment_atlas.xls', sheet = 'RESTAURANTS') %>%
  clean_names() %>%
  select(1:9, 16:17)

store = 
  readxl::read_xls('../food_enviroment_atlas.xls', sheet = 'STORES') %>%
  clean_names() %>%
  select(1:27)

#diabetes rate of each state in 2009
diab2009 =  readr::read_csv('../diabetes2009.csv') %>%
  clean_names() %>% 
  select(1:2) %>% 
  mutate(state = state.abb[match(state,state.name)],
         year = "2009")
diab2009$state[9] = "DC"

#diabetes rate of each state in 2014
diab2014 =  readr::read_csv('../diabetes2014.csv') %>%
  clean_names() %>% 
  select(1:2) %>% 
  mutate(state = state.abb[match(state,state.name)],
         year = "2014")
diab2014$state[9] = "DC"

medhincome_08 = readxl::read_xls("../mhi_08.xls", range = "A13:B63",col_names = c("state","medhhinc08")) %>%
  clean_names() %>%
  mutate(state = state.abb[match(state,state.name)])
medhincome_08$state[17] = "DC"

medhincome_13 = readxl::read_xls("../mhi_13.xls", range = "A11:B61",col_names = c("state","medhhinc13")) %>%
  clean_names() %>%
  mutate(state = state.abb[match(state,state.name)])
medhincome_13$state[8] = "DC"
medhincome_13$`income level` = 
  ifelse(medhincome_13$medhhinc13 < 47242.68,"below average",
         ifelse(medhincome_13$medhhinc13 < 58380.28, "middle income", "above average"))

medhincome = merge(medhincome_08, medhincome_13,by=c("state"))
```
 
## Results
We first provide a overview of the data about diabetes and obesity, then we link those health related data to other possible predictors, such as socioeconomic status, number of fast-food restaurants, National School Lunch Program (NSLP)/School Breakfast Program (SBP) and so on.

### National diabetes and obesity data in 2008 and 2013
We first take a look at the distribution of diabetes and obesity in USA in 2013. Deeper color indicates higher rates of diabetes or obesity. Southeastern part has higher diabetes and obesity rates.  
```{r map}
health_map = 
  health %>%
  group_by(state) %>%
  summarise(diabetes = mean(pct_diabetes_adults13),
            obesity = mean(pct_obese_adults13)) %>%
  ungroup() %>%
  na.omit() %>%
  mutate(full_state = state.name[match(state, state.abb)],
         hover = with(., paste(full_state, '<br>')))

geo_info <- list(
  scope = 'usa',
  projection = list(type = 'albers usa'),
  showlakes = TRUE,
  lakecolor = toRGB('white')
)

plot_geo(health_map, locationmode = 'USA-states') %>%
add_trace(
  z = ~diabetes, text = ~hover, locations = ~state,
  color = ~diabetes, colors = 'OrRd'
) %>%
colorbar(title = "Diabetes Rate (%)") %>%
layout(
  title = 'Adult Diabetes Rate, 2013',
  geo = geo_info
  )

plot_geo(health_map, locationmode = 'USA-states') %>%
add_trace(
  z = ~obesity, text = ~hover, locations = ~state,
  color = ~obesity, colors = 'YlGnBu'
) %>%
colorbar(title = "Obesity Rate (%)") %>%
layout(
  title = 'Adult Obesity Rate, 2013',
  geo = geo_info
  )
```
We then take a look at the distribution of diabetes and obesity nationally.Paired t-test results indicates that both diabetes and obesity rate significantlly increase from 2008 to 2013. This conclusion is consist with our literature research. Diabtes and obesity rates keep rising in the past decades in USA.(Mokdad A H, et al., 2001)
```{r diabetes_obesities_overyear, warning = FALSE, message = FALSE}
require(gridExtra)
t_test_diabetes = 
  health %>%
  ggplot() +
  geom_density(aes(x = pct_obese_adults08), color = "#56B4E9", fill ="#56B4E9", alpha = .4) +
  geom_density(aes(x = pct_obese_adults13), color = "#D55E00",fill ="#D55E00", alpha = .4) +
  labs(title = 'Obesity rate in 08 and 13',
       x = 'rate of obesity')

t_test_obesity = 
  health %>%
  ggplot() +
  geom_density(aes(x = pct_diabetes_adults08), color = "#56B4E9", fill ="#56B4E9", alpha = .4) +
  geom_density(aes(x = pct_diabetes_adults13), color = "#D55E00",fill ="#D55E00", alpha = .4) +
  labs(title = 'Diabetes rate in 08 and 13',
       x = 'rate of diabetes')

grid.arrange(t_test_diabetes, t_test_obesity, ncol=2)

t.test(health$pct_obese_adults08, health$pct_obese_adults13, paired = TRUE) 
t.test(health$pct_diabetes_adults08, health$pct_diabetes_adults13, paired = TRUE)
```
Finding the similar distribution of diabetes and obesity, we continue to test the relationship between diabetes and obesity in 2008 and 2013 respectively. After the linear regression, we find that there are significant linear relationship between diabetes and obesity both in 2008 and 2013. Around 50% variation of diabetes could be explained by obesity. This conclusion is consistent with other national researches.(Nguyen, et al., 2011)
```{r relationship_dibetes_obesity, message = FALSE, warning = FALSE}
require(gridExtra)

relation_2008 = 
  health %>%
  ggplot() +
  geom_point(aes(x = pct_obese_adults08, y = pct_diabetes_adults08), color="#56B4E9", alpha = 0.5)+
  geom_smooth(aes(x = pct_obese_adults08, y = pct_diabetes_adults08), color="#0072B2", method = 'lm') +
  labs(title = 'Relation in 2008',
       x = 'obesity percentage',
       y = 'diabetes percentage')

relation_2013 = 
  health %>%
  ggplot() +
  geom_point(aes(x = pct_obese_adults13, y = pct_diabetes_adults13), color="#F78725", alpha = 0.5)+
  geom_smooth(aes(x = pct_obese_adults13, y = pct_diabetes_adults13), color="#D55E00", method = 'lm') +
  labs(title = 'Relation in 2013',
       x = 'obesity percentage',
       y = 'diabetes percentage')
grid.arrange(relation_2008, relation_2013, ncol=2)

lm(pct_diabetes_adults08~pct_obese_adults08, data = health) %>% summary()
lm(pct_diabetes_adults13~pct_obese_adults13, data = health) %>% summary()
```

### The relationship between diabetes rate and socialeconomics status
```{r scioeconomic_diabetes}
diabetes = health %>%
  group_by(state) %>%
  summarise(pct_08 = mean(pct_diabetes_adults08),
            pct_13 = mean(pct_diabetes_adults13)) %>%
  gather(key = year, value = pct_diabetes_adults, pct_08:pct_13)%>%
  separate(year, into = c("remove", "year"), sep = "_") %>%
  select(-remove)

obese = health %>%
  group_by(state) %>%
  summarise(pct_08 = mean(pct_obese_adults08),
            pct_13 = mean(pct_obese_adults13)) %>%
  gather(key = year, value = pct_obese_adults, pct_08:pct_13)%>%
  separate(year, into = c("remove", "year"), sep = "_") %>%
  select(-remove)

health_status = merge(diabetes, obese, by=c("state", "year"))

income = medhincome %>%
  select(-`income level`) %>%
  gather(key = year, value = medhhinc, medhhinc08:medhhinc13)%>%
  separate(year, into = c("remove", "year"), sep = -3) %>%
  select(-remove)

eco_health = merge(health_status, income, by=c("state", "year"))

eco_health %>%
  group_by(year) %>%
  ggplot(aes(x = medhhinc, y = pct_diabetes_adults)) +
  geom_point(aes(color = year), size = 3, alpha = .8) +
  geom_smooth(se = FALSE) + 
  facet_grid(. ~ year) +
  labs(title = "Median household income and adult diabetes rate relationship",
    x = "Median household income(Dollars)",
    y = "Percentage of adult diabetes")  

summary(eco_health %>% 
          filter(year == "13", medhhinc < 60000) %>%
          lm(pct_diabetes_adults ~ medhhinc, data = .))

summary(eco_health %>% 
          filter(year == "13", medhhinc >= 60000) %>%
          lm(pct_diabetes_adults ~ medhhinc, data = .))
```

Scatterplots are made to show the association between median household income and adult diabetes rate in the U.S. in both 2008 and 2013.  Overall, we can observe that  the adult diabetes rate increased from 08 to 13. Comparing the two plots, we can observe a similar pattern: when the median household income is below 60,000 dollars, there is a decreasing trend; when the median household income is above 60,000 dollars, there is an increasing trend. 

Two separate linear models are fitted regarding the 60,000 dollars threshold for 2013??s data.  The one for below 60,000 dollars is proved to be significant, with a $R^{2}$ of 0. 58, indicating that 58% variation in the adult diabetes rate can be explained by the median household income. However, there is not statistical significant linear relationship between diabetes rate and median household income when the income is above 60,000 dollars.


### The relationship between diabetes rate and restaurant
We would like to find out any association between obesity & diabetes and the information given by the `restaurant` sheet in our dataset.  

First, we tidy and combine our sheets and make the data present in the terms of state. The variables include: Adult diabetes rate in 2013, Adult Obesity rate in 2013, the number of Fast-food restaurants in 2014, the number of Fast-food restaurants per 1,000 population in  2014 and the Expenditures per capita of fast food in 2012.

```{r restaurant_cleaning}
compare = left_join(health, restaurant, by="fips")

compare1 = compare %>%
  group_by(state.x) %>%
  summarize(pct_diabetes_adults13 = mean(pct_diabetes_adults13, na.rm = TRUE),
            pct_obese_adults13 = mean(pct_obese_adults13, na.rm = TRUE),
            ffr14 = sum(ffr14),
            ffrpth14 = mean(ffrpth14),
            pc_ffrsales12 = mean(pc_ffrsales12, na.rm = TRUE))
```

Then, we would like to fit a multiple linear regression model. Considering the fact that we have several predictors, we believe that the stepwise regression would give us the best result. The response variable would be the `Adult diabetes rate in 2013`.

After running stepwise regression, the remaining variable were `the number of Fast-food restaurants per 1,000 population in  2014` and the `Expenditures per capita of fast food in 2012`. The detailed is shown in the `code` part.

```{r restaurant_lm}
mult.fit <- lm(pct_diabetes_adults13 ~ ffr14 + ffrpth14 + pc_ffrsales12, data=compare1)
step(mult.fit, direction='both')

lm(pct_diabetes_adults13 ~ ffrpth14 + pc_ffrsales12, data=compare1) %>% summary()
```

From the summary, we found that the global as well as the two parameters' p-value is less than 0.05 and the adjusted r-squared value is approximately 0.11. Therefore, we believe that there's a significant linear relationship. Unfortunately, the relation is not strong.

Thus, the best fitted model would be: 
$$Y_{i, diabetes rate} = 10.35 - 5.27 \cdot X_{ffrpth14} + 0.006 \cdot X_{pcffrsales12}$$

Scatter plot for this model is below:
```{r restaurant_scatter}
p <- plot_ly(compare1, x = ~ffrpth14, y = ~pc_ffrsales12, z = ~pct_diabetes_adults13, size = ~pct_diabetes_adults13, sizes = c(50, 200),color = ~pct_diabetes_adults13, colors = c('#BF382A', '#0C4B8E')) %>%
  add_markers(text = compare1$state.x) %>%
  layout(scene = list(xaxis = list(title = 'Restaurants per 1000 population'),
                     yaxis = list(title = 'Expenditures per capita'),
                     zaxis = list(title = 'adults diabetes rates in 2013')))
p
```

### The relationship between diabetes rate and school breakfast/lunch program participant propotion

In the following section, these two relationships are analyzed:

* The relationship between diabetes rate (2013) and average School Breakfast program (SBP) participation rate(%) (average between 2009 and 2015).
* The relationship between diabetes rate (2013) and National School Lunch Program (NSLP) participation rate(%) (average between 2009 and 2015).
```{r diabetes_SBP_relation}
participation_rate = assistance %>% 
  group_by(state) %>% 
  mutate(`School Lunch Program` = mean(pct_nslp09, na.rm = T) + mean(pct_nslp15, na.rm = T)) %>% 
  mutate(`School Breakfast Program` = mean(pct_sbp09, na.rm = T) + mean(pct_sbp15, na.rm = T)) %>% 
  ungroup(state) %>% 
  select(state,`School Lunch Program`,`School Breakfast Program`) %>% 
  filter(!(duplicated(state)))

health_13 = health %>% 
  group_by(state) %>% 
  mutate(diabetes_13 = mean(pct_diabetes_adults13, na.rm = T),
         obesites_13 = mean(pct_obese_adults13, na.rm = T)) %>% 
  mutate(statefips = str_sub(fips, 1, 2))%>% 
  filter(!duplicated(state)) %>% 
  ungroup(state) %>%
  select(state, diabetes_13, obesites_13) %>% 
  left_join(participation_rate, by = "state") %>% 
  left_join(medhincome_13, by = "state") %>% 
  gather(key = "program", value = "participant proportion", `School Lunch Program`:`School Breakfast Program`)

ggplot(health_13,aes(x = `participant proportion`, y = diabetes_13)) +
  geom_point(aes(color = `income level`)) +
  geom_smooth(method = "lm",formula = y~x) +
     ggpmisc::stat_poly_eq(formula = y~x,
            aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~")), 
            parse = TRUE) +
  labs(title = "Diabetes Rate vs. Program Participant",
    x = "Participant Proportion(%)",
    y = "Diabetes Rate(%)") +
  theme(legend.position = "bottom") +
  facet_grid(. ~program, scales = "free")
```  

**Note:** the `Income level` refers to the level median household income of each state. `Above average` income level is defined as median household income above 75% quantile of all the states. `Below average` income level is defined as median household income below 25% quantile of all the states. `Middle` income is between 25% quantile and 75% quantile.

The linear relationship between diabetes rate (2013) and average SBP participant proportion or average NSLP participant proportion are both significant with adjusted $R^{2}$ of 40% and 26% respectively. The results indicate that there is a positive relationship. However, it may be confounded by social and economic status.

From the scatter plot, we can see that most states with below average median household income (green dots) are above the regression line and most states with above average household income (red dots) are below the line. Therefore, income level may differ the relationship between participant proportion and diabetes rate. In those states with lower average household income, for every unit increase in participant proportion, the expected increase in diabetes rate may be higher than states with higher household income.

### The relationship between diabetes rate and the average supermarkets & grocery stores number/1,000 residents

Do linear regression between diabetes rate within state and average number of supermarket & grocery store number/1000 residents in each state (average over each county) in 2009 and 2014. 

```{r average_stores}
health_store = store %>% 
  group_by(state) %>% 
  mutate(store_rate_2009 = mean(grocpth09, na.rm = T),
         store_rate_2014 = mean(grocpth14, na.rm = T),
         store_mean = (store_rate_2009+store_rate_2014)/2) %>% 
  select(state,store_rate_2009,store_rate_2014,store_mean) %>% 
  filter(!(duplicated(state)))

health_store1 = health_store %>% 
  select(-store_mean) %>% 
  left_join(medhincome_13, by = "state") %>% 
  gather(key = "year", value = "store_rate", store_rate_2009:store_rate_2014) %>% 
  mutate(year = str_sub(year, -4, -1)) %>% 
  left_join(rbind(diab2009,diab2014), by = c("state","year")) %>% 
  ggplot(aes(x = store_rate, y = percentage)) +
     geom_point(aes(color = `income level`), alpha = 0.7) +
     geom_smooth(method = "lm", formula = y~x) +
     ggpmisc::stat_poly_eq(formula = y~x,
            aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~")), 
            parse = TRUE) +
   coord_cartesian(xlim = c(0.15, 0.5)) +
   labs(title = "",
    x = "Average supermarket & grocery store/1000 residents",
    y = "Diabetes Rate(%)") +
    facet_grid(. ~ year)

health_store1  
```

**Note:** the `Income level` refers to the level median household income of each state. `Above average` income level is defined as median household income above 75% quantile of all the states. `Below average` income level is defined as median household income below 25% quantile of all the states. `Middle` income is between 25% quantile and 75% quantile.

Overall, there is a negative linear relationship between diabetes rate and average number of supermarket & grocery stores/1000 residents with $R^2$ 0.22 and 0.3 respectively. The negative association is consistant with the negative relationship between diabetes rate and food accessibility. Lower accessibility to grocery stores and farmers' market is relateved to higher diabetes and obesitiy rate (Salois, 2012).

States with below average household income are clustered at lower `average number of supermarke & grocery stores/1000 residents` and high `Diabetes Rate`, which is consistant with previous findings. In addition, obesity may be a confounder in the `diabetes rate vs. store number/1000 residents` relationship, since lower food accessibility are both correlated with diabetes and obesity.

### A linear model that capture the overall association
After identifying the food related factors that may have certain relationship with adult diabetes prevalence, we decide to construct an explanatory linear model to capture the association as a whole. Since in previous section, we observed a strong linear association between median household income and adult diabetes rate when the income is below 60,000 dollars, we focus on building a model to explain the association between the covariates identified from previous sections and adult diabetes rate in this sub-population. 

The variables we are interested in to build the model for adult diabetes rate at the state level include: 

* adult obesity rate (obe)
* median household income (MHI)
* lunch program participation rate (LPPR)
* the number of Fast-food restaurants per 1,000 population (FFPP)
* the Expenditures per capita of fast food (EFF)
* the number of grocery store and supermarket per 1,000 population (store)

```{r fit_model}
store_data = health_store %>%
  select(state, store_rate_2014)
lunch_data = health_13[c(1:51),]
colnames(compare1)[1:3] = c("state", "diabetes_13", "obesites_13")
model_data = merge(lunch_data, compare1, by = c("state", "diabetes_13", "obesites_13")) %>%
  merge(.,store_data, by = "state") %>%
  select(-`income level`, - program, - ffr14, - state) %>% 
  filter(medhhinc13 < 60000)

# correlation of variables
correlation_data = cor(model_data)

# find best model
## stepwise and backward
mult.fit.full <- lm(diabetes_13 ~ ., data=model_data)
step(mult.fit.full, direction='both')
step(mult.fit.full, direction='back')

## criterion based
best <- function(model, ...) 
{
  subsets <- leaps::regsubsets(formula(model), model.frame(model), ...)
  subsets <- with(summary(subsets),
                  cbind(p = as.numeric(rownames(which)), which, rss, rsq, adjr2, cp, bic))
  
  return(subsets)
}  

# Select the 'best' 1 model of all subsets
round(best(mult.fit.full, nbest = 1), 4)

best_fit = lm(formula = diabetes_13 ~ obesites_13 + medhhinc13 + `participant proportion` + 
    pc_ffrsales12 + store_rate_2014, data = model_data)

best_fit %>% summary()
```

Based on the result of both stepwise regression, backward elimination (based on AIP), and criterion-based procedures, we get the following model:

$$Y_{diabetes\ rate} = 2.653 + 0.3958 \cdot X_{obe} - 0.00006 \cdot X_{MHI}- 0.06442 \cdot X_{LPPR} + 0.00337 \cdot X_{EFF} - 4.494 \cdot X_{store}$$

```{r assum_influ}
check = influence.measures(best_fit)
# par(mar=c(4,4,1,1))
# par(mfrow=c(2,2))
# plot(best_fit)
```

After checking the model assumptions and influential values, we conclude that this model is a good one to capture the association. The adjust R square is 0.885,
Indicating that 88.5% variation in adult obesity rate can be explained by this linear model. 

## Discussion
At the beginning of this project, we take a look at the diabetes and obesity data in 2008 and 2013. We find that there are strong relationship between diabetes and obesity, meanwhile, southeastern US has higher diabetes and obesity rates. Combining our tests and literature research, it is worth mentioning that national diabetes and obesity rates increase over decades.

We then try to compare the diabetes rate with other possible covariables. When the median household income is below 60,000 dollars, there is a decreasing trend between income and diabetes rate, while when the income is above, 60,000 dollars, the trend turns to be positive. As for the restaurant data, combining the number of restaurants per 1000 population with the expenditures per capita on fast food, there??s a linear relationship with diabetes rates. Surprisingly, there is a positive relationship between diabetes rate and National School Lunch Program/School Breakfast Program participation proportion, which is counterintuitive -- NSLP and SBP serve nutritious food to children and should negatively correlated with diabetes rate It may be because the association is confounded by other factors such as social-economic factors because children who are eligible for NSLP and SBP tend to come from families with low income. 






## Reference
* Centers for Disease Control and Prevention. "[National diabetes statistics report: estimates of diabetes and its burden in the United States, 2017.](https://www.cdc.gov/diabetes/pdfs/data/statistics/national-diabetes-statistics-report.pdf)" Atlanta, GA: National Center for Chronic Disease Prevention and Health Promotion .

* Mathers, Colin D., and Dejan Loncar. "[Projections of global mortality and burden of disease from 2002 to 2030.](http://journals.plos.org/plosmedicine/article?id=10.1371/journal.pmed.0030442)" PLoS medicine 3.11 (2006): e442.

* Mokdad, Ali H., et al. "The continuing epidemics of obesity and diabetes in the United States." Jama 286.10 (2001): 1195-1200.

* Nguyen, Ninh T., et al. "[Relationship between obesity and diabetes in a US adult population: findings from the National Health and Nutrition Examination Survey, 1999-2006.](https://link.springer.com/content/pdf/10.1007%2Fs11695-010-0335-4.pdf)" Obesity surgery 21.3 (2011): 351-355.

* Salois, Matthew J. "Obesity and diabetes, the built environment, and the ???local???food economy in the United States, 2007." Economics & Human Biology 10.1 (2012): 35-42.

* Smyth, Simon, and Andrew Heron. "[Diabetes and obesity: the twin epidemics.](https://www.nature.com/articles/nm0106-75)" Nature medicine 12.1 (2006): 75-80.

* World Health Organization. [Global report on diabetes](http://apps.who.int/iris/bitstream/10665/204871/1/9789241565257_eng.pdf?ua=1). World Health Organization, 2016.  



